Human Endogenous Retroviral Transcripts Show Potential as Novel, Tumor Specific Biomarkers for Colorectal Cancer

### 2021-01-24 manuscript/secs/results.tex
add supp table with HERV coordinates

To verify the consistency of associations between proviral expression and tissue phenotype and to validate the utility of tumor specific HERVs as potential biomarkers, we applied the same methods used in the discovery cohort to an independent cohort (cohort \#1) of 276 colorectal samples composed of 12 healthy, 56 tumor adjacent, and 208 tumor samples (Supplementary Table 1).
Of the 150 HERVs moderately expressed (i.e. expressed within a standard deviation of median protein coding expression) in the discovery cohort, 92 were also moderately expressed in cohort \#1.
Estimates of relative expression between healthy and tumor samples were similar for the 92 HERVs tested in both cohorts despite the small number of healthy samples in cohort \#1 (Supplementary Figure 1c).
As tumor classifiers in this cohort, most of the 11 tumor specific HERVs had weaker performance (Supplementary Figure 6).
However, HERV-H Xp22.32 and HERV1-I 3p22.3 maintained excellent discriminatory ability.
Seven of the 11 tumor specific HERVs identified in the discovery cohort were validated as consistently tumor specific in cohort \#1 (Table \ref{tab:val}).
Note that nominally significant (i.e. \emph{P} < 0.05) tumor over-expression relative to either healthy or tumor adjacent samples was considered significant for the purpose of validation due to the small number of healthy samples.

>>>>

To verify the consistency of associations between proviral expression and tissue phenotype and to validate the utility of tumor specific HERVs as potential biomarkers, we applied the same methods used in the discovery cohort to an independent cohort (cohort \#1) of 276 colorectal samples composed of 12 healthy, 56 tumor adjacent, and 208 tumor samples (Supplementary Table 1).
Of the 150 HERVs moderately expressed (i.e. expressed within a standard deviation of median protein coding expression) in the discovery cohort, 92 were also moderately expressed in cohort \#1.
Estimates of relative expression between healthy and tumor samples were similar for the 92 HERVs tested in both cohorts despite the small number of healthy samples in cohort \#1 (Supplementary Figure 1c).
As tumor classifiers in this cohort, most of the 11 tumor specific HERVs had weaker performance (Supplementary Figure 6).
However, HERV-H Xp22.32 and HERV1-I 3p22.3 maintained excellent discriminatory ability.
Seven of the 11 tumor specific HERVs identified in the discovery cohort were validated as consistently tumor specific in cohort \#1 (Table \ref{tab:val} and Supplementary Table 3).
Note that nominally significant (i.e. \emph{P} < 0.05) tumor over-expression relative to either healthy or tumor adjacent samples was considered significant for the purpose of validation due to the small number of healthy samples.
